Favourable Conservation Status (FCS): <br> Analysis of Bauer et al. (unpublished) Field experiment

### 2022-11-10 R/model_fcs_target_markdown.Rmd
Separate model building and model check of FCS

>>>>

## Model building

### Models

#### Preparation

```{r}
### Posssible priors ###
get_prior(n ~ target_type + exposition + sand_ratio + survey_year_fct +
            seed_density + substrate_depth +
            (1 | site/plot) + (1 | botanist_year),
          data = sites)
### Example of normal distribution
ggplot(data = data.frame(x = c(-2, 2)), aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0.1, sd = 1))
### Example of cauchy distribution
ggplot(data = data.frame(x = c(-2, 2)), aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1))
### Example of a student t distribution
ggplot(data.frame(x = c(-2, 2)), aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5))
```

#### Calculation

Specifications for the models

```{r}
iter = 10000
chains = 4
thin = 2
priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0.1, 1)", class = "b", coef = "sand_ratio25"),
  set_prior("normal(0.2, 1)", class = "b", coef = "sand_ratio50"),
  set_prior("normal(0.1, 1)", class = "b", coef = "expositionsouth"),
  set_prior("normal(0.1, 1)", class = "b", coef = "survey_year_fct2019"),
  set_prior("normal(0.2, 1)", class = "b", coef = "survey_year_fct2020"),
  set_prior("normal(0.3, 1)", class = "b",coef = "survey_year_fct2021"),
  set_prior("cauchy(0, 1)", class = "sigma")
)
```

Model caluclations

```{r}
m1 <- brm(n ~ (target_type + exposition + sand_ratio + survey_year_fct)^4 +
             substrate_depth + seed_density +
             (1 | site/plot) + (1 | botanist_year),
           data = sites, 
           family = gaussian("identity"),
           prior = priors,
           chains = chains,
           iter = iter,
           thin = thin,
           warmup = floor(iter / 2),
           save_pars = save_pars(all = TRUE),
           cores = parallel::detectCores(),
           seed = 123)

m2 <- brm(n ~ (target_type + exposition + sand_ratio + survey_year_fct)^4 +
            substrate_depth + seed_density +
            sand_ratio:substrate_depth +
            (1 | site/plot) + (1 | botanist_year),
          data = sites, 
          family = gaussian("identity"),
          prior = priors,
          chains = chains,
          iter = iter,
          thin = thin,
          warmup = floor(iter / 2),
          save_pars = save_pars(all = TRUE),
          cores = parallel::detectCores(),
          seed = 123)

m3 <- brm(n ~ (target_type + exposition + sand_ratio + survey_year_fct)^2 +
                 substrate_depth + seed_density +
                 substrate_depth:sand_ratio +
                 seed_density:exposition +
                 target_type:exposition:survey_year_fct +
                 sand_ratio:exposition:survey_year_fct +
                 seed_density:exposition:survey_year_fct +
                 (1 | site/plot) + (1 | botanist_year),
               data = sites, 
               family = gaussian("identity"),
               prior = priors,
               chains = chains,
               iter = iter,
               thin = thin,
               warmup = floor(iter / 2),
               save_pars = save_pars(all = TRUE),
               cores = parallel::detectCores(),
               seed = 123)
```

### Model comparison

```{r}
m_1 <- m1
m_2 <- m3
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

### Model check

#### DHARMa

```{r}
createDHARMa(
  simulatedResponse = t(posterior_predict(m_1)),
  observedResponse = sites$n,
  fittedPredictedResponse = apply(t(posterior_epred(m_1)), 1, mean),
  integerResponse = TRUE
  ) %>%
  plot()
createDHARMa(
  simulatedResponse = t(posterior_predict(m_2)),
  observedResponse = sites$n,
  fittedPredictedResponse = apply(t(posterior_epred(m_2)), 1, mean),
  integerResponse = TRUE
  ) %>%
  plot()
```

#### Preparation

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sd_botanist_year__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sd_botanist_year__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- nuts_params(m_1)
hmc_diagnostics2 <- nuts_params(m_2)
y <- sites$n
yrep1 <- posterior_predict(m_1, draws = 500)
yrep2 <- posterior_predict(m_2, draws = 500)
loo1 <- loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

#### Samling efficency/effectiveness (Rhat and EFF)

```{r}
range(draws1$rhat)
range(draws2$rhat)
range(draws1$ess_bulk)
range(draws2$ess_bulk)
range(draws1$ess_tail)
range(draws2$ess_tail)
```

#### MCMC diagnostics

```{r}
mcmc_trace(posterior1, np = hmc_diagnostics1)
mcmc_trace(posterior2, np = hmc_diagnostics2)
```

#### Posterior predictive check

##### Kernel density

```{r}
ppc_dens_overlay(y, yrep1[1:50, ])
ppc_dens_overlay(y, yrep2[1:50, ])
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$site)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$site)
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$exposition)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$exposition)
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$survey_year_fct)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$survey_year_fct)
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$target_type)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$target_type)
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$seed_density)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$seed_density)
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$sand_ratio)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$sand_ratio)
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$substrate_depth)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$substrate_depth)
```

##### Histograms of statistics skew

```{r}
ppc_stat(y, yrep1, binwidth = 0.001)
ppc_stat(y, yrep2, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
```

##### LOO-PIT plots

```{r}
ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
plot(loo1)
plot(loo2)
```

#### Autocorrelation check

```{r}
mcmc_acf(posterior1, lags = 10)
mcmc_acf(posterior2, lags = 10)
```

## Output of choosen model

### Model output

```{r}
prior_summary(m_1)
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
draws1
mcmc_intervals(
  posterior1,
  prob = 0.5,
  prob_outer = 0.95,
  point_est = "mean"
)
sjPlot::plot_model(m_1, m_2, type = "est")
sjPlot::plot_model(m_2, type = "est")
sjPlot::plot_model(
  m_1, m_2, type = "est", ppd = TRUE,
  terms = c(
    "sand_ratio", "substrate_depth", "target_type",
    "seed_density",  "exposition", "survey_year_fct"
    )
  )
```

### Effect sizes

```{r}
(emm <- emmeans(m_1, revpairwise ~ target_type + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
(emm <- emmeans(m_1, revpairwise ~ substrate_depth + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
(emm <- emmeans(m_1, revpairwise ~ seed_density |
                  exposition | survey_year_fct, type = "response"))
```
